\section{Bibliography}

\begin{enumerate}[label={[\arabic*]}]
    \item Bragarenco, A., Astafi, V., \textit{Sisteme Electronice Încorporate: Indicații metodice pentru lucrări de laborator}, Universitatea Tehnică a Moldovei, Chișinău, 2024.

    \item Technical University of Moldova, \textit{Embedded Systems: Practical Guide for Hardware-Software Interface Development}, Department of Microelectronics and Biomedical Engineering, Chișinău, 2024.

    \item Maxim Integrated, \textit{DS18B20 Programmable Resolution 1-Wire Digital Thermometer}, Datasheet Rev. 042208, 2019. Available: \url{https://datasheets.maximintegrated.com/en/ds/DS18B20.pdf}

    \item Murata Manufacturing Co., \textit{NTC Thermistors --- NTC Thermistor Technical Note}, Application Note, 2021. Available: \url{https://www.murata.com/en-us/support/faqs/products/thermistor/ntc}

    \item Richard Barry, \textit{Mastering the FreeRTOS Real Time Kernel --- A Hands-On Tutorial Guide}, Real Time Engineers Ltd., 2016. Available: \url{https://www.freertos.org/Documentation/RTOS_book.html}

    \item Arduino, \textit{Arduino Mega 2560 Rev3 Documentation}, \url{https://docs.arduino.cc/hardware/mega-2560/}, Accessed: 2025-04-10.

    \item PlatformIO, \textit{Professional collaborative platform for embedded development}, \url{https://platformio.org/}, Accessed: 2025-04-10.

    \item Wokwi, \textit{Online Electronics Simulator --- Arduino, ESP32, Raspberry Pi Pico}, \url{https://wokwi.com/}, Accessed: 2025-04-10.

    \item Miles Burton, \textit{DallasTemperature Arduino Library}, \url{https://github.com/milesburton/Arduino-Temperature-Control-Library}, Accessed: 2025-04-10.

    \item Paul Stoffregen, \textit{OneWire Arduino Library}, \url{https://github.com/PaulStoffregen/OneWire}, Accessed: 2025-04-10.
\end{enumerate}
