\section{Note Regarding Usage of AI}

During the preparation of this laboratory report, the author utilized artificial intelligence tools to enhance the quality and presentation of the documentation. Specifically, the following AI tools were employed:

\begin{itemize}
    \item \textbf{Report Generation:} Claude Code (Anthropic) was used to generate the full LaTeX report structure, all section content, TikZ diagrams, and CircuiTikZ schematics based on the implemented code and architecture.
    \item \textbf{Code Implementation:} Claude Code assisted in generating the \texttt{BinaryConditioner} and \texttt{ActuatorControl} library modules, the lab4\_1 FreeRTOS task files, and the Wokwi simulation configuration.
    \item \textbf{Report Formatting:} Assistance with LaTeX document structure, formatting consistency, and adherence to the established laboratory report template.
    \item \textbf{Code Documentation:} Generation of inline Doxygen-style comments within the source code to improve readability and maintainability.
\end{itemize}

All AI-generated content was thoroughly reviewed, validated, and adjusted by the author to ensure technical accuracy, alignment with project requirements, and adherence to embedded systems best practices. The core technical architecture decisions, signal conditioning pipeline design, FreeRTOS task structure, and system integration represent the author's understanding of the subject matter. The AI tools served as assistive technologies to enhance documentation quality and accelerate the development workflow, while the author retained full responsibility for the content, correctness, and conclusions presented in this report.
