\section{Note Regarding Usage of AI}

During the preparation of this laboratory report, the author utilized artificial intelligence tools to enhance the quality and presentation of the documentation. Specifically, the following AI tools were employed:

% TODO: Update the list of AI tools and their usage as applicable.

\begin{itemize}
    \item \textbf{Report Formatting:} Assistance with LaTeX document structure, formatting consistency, and adherence to academic writing standards.
    \item \textbf{Report Template Generation:} Creation of the initial document template with proper sectioning and organizational structure.
    \item \textbf{Content Rephrasing:} Refinement of technical descriptions and explanations to ensure grammatical accuracy and clarity without altering the technical substance.
    \item \textbf{Code Documentation:} Generation of inline comments within the source code to improve code readability and maintainability.
    \item \textbf{Development Environment Setup:} Configuration assistance for PlatformIO integration with the Wokwi simulation platform.
\end{itemize}

All AI-generated content was thoroughly reviewed, validated, and adjusted by the author to ensure technical accuracy, alignment with project requirements, and adherence to embedded systems best practices. The core technical implementation, circuit design, system architecture decisions, and experimental results represent the author's original work and understanding of the subject matter. The AI tools served as assistive technologies to enhance documentation quality and development workflow efficiency, while the author retained full responsibility for the content, correctness, and conclusions presented in this report.
